Peaks Over Threshold for Bursty Time Series

### 2018-04-18 article_springer/article_springer.Rmd
pick 2 of the three plots

Figures \ref{fig:flare-diagnostics-1} and \ref{fig:flare-diagnostics-2} 
show the diagnostic plots for a 
minimum threshold chosen at the 200th order statistic. 
There is some residual autocorrelation for the sequence of threshold 
exceedance times that is not accounted for by the MRP model. 
The fit with a Mittag-Leffler distribution ($\beta = 0.8$) is good, and 
there is no apparent dependence between threshold exceedance times and 
event magnitudes.

>>>>

Figures \ref{fig:flare-diagnostics-1} and \ref{fig:flare-diagnostics-2} 
show the diagnostic plots for a 
minimum threshold chosen at the 200th order statistic. 
There is some residual autocorrelation for the sequence of threshold 
exceedance times that is not accounted for by the MRP model. 
The fit with a Mittag-Leffler distribution ($\beta = 0.8$) is good, 
though there are signs that the power-law tail tapers off for very large 
inter-threshold crossing times. 
There is no apparent dependence between threshold exceedance times and 
event magnitudes seen in the copula plot.

# Predicting the time of the next threshold crossing

According to Figure \ref{fig:flares}, for a threshold $\ell$ at the $k$-th order 
statistic, the fitted threshold exceedance time distribution is 
$$
T_\ell \sim {\rm ML}(\beta, k^{1/\beta} \sigma_0), 
$$
where $\beta = 0.85$ and $\sigma_0 = 3.0 \times 10^7 {\rm sec}$. 
Unlike the exponential distribution, the Mittag-Leffler distribution is not 
memoryless, and the probability density of the time $t$ until the next threshold 
crossing will depend on the time $t_0$ elapsed since the last threshold crossing.
This density equals 
$$
p(t|\beta, \sigma_0, \ell, t_0) = \frac{f(t + t_0 | \beta, k^{1/\beta} \sigma_0)}{\mathbf P[T_\ell > t_0]}
$$
where $f(\,\cdot\, | \beta, k^{1/\beta} \sigma_0)$ is the probability density of 
${\rm ML}(\beta, k^{1/\beta} \sigma_0)$. 
The more time has passed without a threshold crossing, the more the probability 
distribution shifts towards larger values for the next crossing
(see Figure \ref{fig:hazard}, left panel). 
The hazard rate 
$$
h(t) = \frac{f(t| \beta, k^{1/\beta} \sigma_0))}{\int_t^\infty f(\tau| \beta, k^{1/\beta} \sigma_0))\,d\tau}
$$
represents the risk of a threshold crossing per unit time, and is a 
decreasing function for the Mittag-Leffler distribution. 
The closer $\beta$ is to $1$, the more the hazard rate mimics that of an 
exponential distribution (a constant function, see Figure \ref{fig:hazard}, 
right panel).

Conditional on the last threshold being

```{r hazard, fig.height=7, fig.cap="Left: Conditional distribution of time until next threshold crossing, depending on elapsed time since last crossing ($\\beta = 0.8$, $\\sigma_0 = 1$). Right: Hazard rate depending on tail parameter $\\beta$.\\label{fig:hazard}"}
tail <- 0.7
scale <- 1
t_0 <- c(0, 1, 10)
from <- 0.1
to <- 100 * scale
xx <- exp(seq(from = log(from), to = log(to), length.out = 100))
par(mfrow = c(2,2))
plot(xx, rep(1, length(xx)), ylim = c(0.001,1), type = 'n', log = 'xy', 
     xlab = 't', ylab = 'p')
for (k in 1:length(t_0)) {
  cond <- pml(q = t_0[k], tail = tail, scale = scale, lower.tail = FALSE)
  yy <- dml(x = xx + t_0[k], tail = tail, scale = scale) / cond
  lines(xx,yy, col = k+1, lty = k)
}
legend("bottomleft", c("t_0 = 0", "t_0 = 1", "t_0 = 10"), col = 2:4, lty = 1:3)

betas <- c(0.7, 0.8, 0.95, 0.99)
t_0 <- 1
plot(xx, rep(1, length(xx)), ylim = c(0.001,2), type = 'n', log = 'xy', 
     xlab = 't', ylab = 'h')
for (k in 1:length(betas)) {
  yy <- dml(x = xx, tail = betas[k], scale = scale) / pml(
    q = xx,
    tail = betas[k],
    scale = scale,
    lower.tail = FALSE
  )
  lines(xx,yy, col = k+1, lty = k)
}
legend("bottomleft", c("beta = 0.7", "beta = 0.8", "beta = 0.95", "beta = 0.99"), col = 2:5, lty = 1:4)

```
